Localisation and Sensor Privacy Using the Extended Information Filter and Private Linear-Combination Aggregation\thanksref{footnoteinfo}

### 2020-10-12 main.tex
Added references for the placeholder citations - removed a couple as well

\section{Introduction}
Localisation methods in distributed sensor environments have long been an active topic of research []. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localisation methods, from range-free and range-based signal strength measurements [], echo-based range estimates [] and magnetic field measurements [] to the TDOA and TOA pseudo-range multilateration methods []. Range-based localisation methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large application due to their robustness and accuracy []. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally []. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made these requirements of information security and user privacy more apparent than before [14,15]. For example, a scenario of aircraft localisation in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Alternatively, autonomous vehicle state estimation may benefit from neighbouring vehicles' information while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localisation filter using this scheme such that sensor and navigator privacy is preserved.

@@

Of the methods used to preserve privacy during distributed signal processing, it is common for the information exchanging portion of the algorithm to be reduced to a form of aggregation of involved parties' data []. The most common methods for ensuring privacy during aggregation have been described below.

@@

    \item[Differential Privacy] These techniques formalise preserving the privacy of individual datum within an aggregation operation by introducing known distributions of noise to the resulting aggregation []. In the cases with no trusted aggregator, local differential privacy is achieved by introducing noise to individual inputs [norm, PHE ones]. Although differential privacy has strong computational benefits due to its implementational simplicity, it provides relatively weak security guarantees for multi-party networks, with no guarantees in the presence of multiple colluding corrupted parties. Additionally, noisy aggregation results hinder its use in scenarios where high precision is a desired property such as aircraft localisation.
    \item[Aggregator Oblivious Encryption] This method of encryption formalises the privacy of \textit{all} aggregation data from honest parties []. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method [] is restricted to a small input domain which has been expanded and computationally simplified in later works [joye-lib,student-one]. The aggregation operation in aggregator obliviousness is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter mean calculation [].

@@

    \item[Fully Homomorphic Encryption] These schemes allow all algebraic computations to be performed on encrypted data without the computing party learning any information about inputs, intermediate values or results []. Although theoretically ideal for many privacy-preserving tasks, current implementations are still computationally impractical for real-time or large scale tasks [].
    \item[Partially Homomorphic Encryption] A simplification of fully homomorphic schemes, these schemes provide only a subset, typically one, of algebraic operations on encrypted data []. The reduced computational requirements and implementation simplicity have made partially homomorphic encryption schemes commonly used in privacy-preserving signal processing tasks, including private matrix multiplication [], set intersection computation [] and control input aggregation []. However, these works are relatively restricted in application due to the limited operations provided.

@@

    \item[Multi-Client Function Encryption] This type of encryption allows the computing of a specific function given encrypted inputs []. It is suitable for privacy-preserving distributed data as encryptions must be combined before a result is obtained, and can be considered a generalisation of the aggregation schemes described earlier. While such schemes that can compute arbitrary functions are still too computationally expensive for practical use [], partial schemes supporting only some functions, such as computing vector dot products [], exist, but have yet to find applications in more complicated signal processing tasks.

@@

\subsection{Encrypted Localisation and Estimation}
Localisation methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors []. We aim to provide an accurate localisation filter in the presence of range-only sensors.

Relevant estimation methods include both model-based Bayesian estimation methods which make assumptions about target trajectories to form their location estimates, typically based on the Kalman Filter [], and model-free methods where location estimates are made independently of each other [proloc,serloc]. The work in [proloc] uses partially homomorphic encryption to compute model-free location estimates while preserving the privacy of sensor measurements and locations. The work proposes two approaches, using either polygon intersections suffering from geometric dilution of precision, or alternating projection requiring repeated interactive comparison protocols. Neither method considers measurement uncertainty, and no estimate errors are provided.

Model-based approaches can achieve better accuracy due to their exploitation of process dynamic knowledge. In [aristov], partially homomorphic encryption is used to encrypt measurement information and produce location estimates at a navigator. While producing estimate errors and requiring only uni-directional communication, this method exclusively supports linear measurement models and is therefore not suitable for range-only measurements. Model-based encrypted control [farohki summary] also introduces some privacy-preserving methods relevant to estimation. [alexandru] proposes secure distributed control input aggregation using a novel weighted sum aggregation scheme, but requires repeated key distribution among all involved parties. In [farohki], a secure cloud-based control aggregator using partially homomorphic encryption is proposed, but is suitable only for untrusted service-providing cloud architectures.

While we primarily focus on applying encryption schemes to localisation algorithms, homomorphic encryption schemes also require suitable number encoding when used with real numbers. As is the case with localisation sensors, real number measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. The Google BigNum library [google bignum] provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In [farohki], a simpler alternative for encoding is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings.

@@

    Misbehaving sensors are a known problem in estimation theory, often requiring inconsistency or outlier detection mechanisms. As privacy-preserving implementations of such mechanisms require additional complications, we do not consider the scenario in this work.
    \item[Computational capabilities] The computational requirements for computing encryptions and homomorphic operations are typically more costly than their plaintext equivalents. We make the implicit assumption that all involved parties are computationally capable of computing the required encryption and filter operations.
\end{description}
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\footnotesize]
    % Sizes
    \tikzmath{\srange = 3.54; \sradius = 0.04;}
    % Sensor locations (sensor 3 is shared)
    \coordinate (s1) at (0.8523,-4.1508) {};
    \coordinate (s2) at (3.507,-5.5587) {};
    \coordinate (s3) at (2.5,-2.75) {} {};
    \coordinate (s4) at (4,-1.25) {};
    \coordinate (s5) at (5.4346,-3.1274) {};
    % Navigator location
    \coordinate (nav_loc) at (3.5,-3.25) {};
    % Bounding rectangle
    \coordinate (brect_start) at (0,0);
    \coordinate (brect_end) at (6.5, -6.5);

@@

% Include this if you use bibtex 
% and a bib file to produce the 
% bibliography (preferred). The
% correct style is generated by
% Elsevier at the time of printing.
\bibliographystyle{plain}
\bibliography{BibTeX/PrivateEIFLocalisation}

>>>>

\section{Introduction}
Localisation methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localisation methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004} and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localisation methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large application due to their robustness and accuracy. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made these requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}. For example, a scenario of aircraft localisation in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Alternatively, autonomous vehicle state estimation may benefit from neighbouring vehicles' information while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localisation filter using this scheme such that sensor and navigator privacy is preserved.

@@

Of the methods used to preserve privacy during distributed signal processing, it is common for the information exchanging portion of the algorithm to be reduced to a form of aggregation of involved parties' data. The most common methods for ensuring privacy during aggregation required for signal processing tasks have been described below.

@@

    \item[Differential Privacy] These techniques formalise preserving the privacy of individual datum within an aggregation operation by introducing known distributions of noise to the resulting aggregation. To handle the cases with no trusted aggregator, local differential privacy is achieved by introducing noise to individual inputs \cite{hanPrivacyControlDynamical2018,dworkDifferentialPrivacySurvey2008,andresGeoindistinguishabilityDifferentialPrivacy2013}. Although differential privacy has strong computational benefits due to its implementational simplicity, it provides relatively weak security guarantees for multi-party networks, with no guarantees in the presence of multiple colluding corrupted parties. Additionally, noisy aggregation results hinder its use in scenarios where high precision is a desired property such as aircraft localisation.
    \item[Aggregator Oblivious Encryption] This method of encryption formalises the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in aggregator obliviousness is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter mean calculation \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.

@@

    \item[Fully Homomorphic Encryption] These schemes allow all algebraic computations to be performed on encrypted data without the computing party learning any information about inputs, intermediate values or results \cite{gentryFullyHomomorphicEncryption2009,stehleFasterFullyHomomorphic2010}. Although theoretically ideal for many privacy-preserving tasks, current implementations are still computationally impractical for real-time or large scale tasks \cite{acarSurveyHomomorphicEncryption2018}.
    \item[Partially Homomorphic Encryption] A simplification of fully homomorphic schemes; these schemes provide only a subset, typically one, of algebraic operations on encrypted data \cite{elgamalPublicKeyCryptosystem1985,paillierPublicKeyCryptosystemsBased1999,bonehEvaluating2DNFFormulas2005}. Reduced computational requirements and implementation simplicity have made partially homomorphic encryption schemes commonly used in privacy-preserving signal processing tasks, including private matrix multiplication \cite{kogisoCyberSecurityEnhancementNetworked2015}, set intersection computation \cite{kerschbaumOutsourcedPrivateSet2012} and control input aggregation \cite{alexandruEncryptedCooperativeControl2019}. However, these works are relatively restricted in application due to the limited operations provided.

@@

    \item[Multi-Client Function Encryption] These schemes allow the computing of a specific function given encrypted inputs \cite{bonehFunctionalEncryptionDefinitions2011,goldwasserMultiinputFunctionalEncryption2014}. It is suitable for privacy-preserving distributed data as encryptions must be combined before a result is obtained, and can be considered a generalisation of the aggregation schemes described earlier. While general schemes that allow computing arbitrary functions are computationally expensive and complicate security definitions in distributed settings \cite{agrawalFunctionalEncryptionNew2013}, partial schemes supporting only some functions, such as vector dot products \cite{chotardDecentralizedMultiClientFunctional2018}, exist with formal security proofs, but have yet to find applications in more complicated signal processing tasks.

@@

\subsection{Encrypted Localisation and Estimation}
Localisation methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors. We aim to provide an accurate localisation filter in the presence of range-only sensors.

Relevant estimation methods include both model-based Bayesian estimation methods which make assumptions about target trajectories to form their location estimates, typically based on the Kalman Filter and later derivatives \cite{kalmanNewApproachLinear1960,mutambaraDecentralizedEstimationControl1998}, and model-free methods where location estimates are made independently of each other \cite{alanwarPrOLocResilientLocalization2017,lazosSeRLocSecureRangeindependent2004}. The work in \cite{alanwarPrOLocResilientLocalization2017} uses partially homomorphic encryption to compute model-free location estimates while preserving the privacy of sensor measurements and locations. The work proposes two approaches, using either polygon intersections suffering from geometric dilution of precision, or alternating projection requiring repeated interactive comparison protocols. Neither method considers measurement uncertainty, and no estimate errors are provided.

Model-based approaches can achieve better accuracy due to their exploitation of process dynamic knowledge. In \cite{aristovEncryptedMultisensorInformation2018}, partially homomorphic encryption is used to encrypt measurement information and produce location estimates at a navigator. While producing estimate errors and requiring only uni-directional communication, this method exclusively supports linear measurement models and is therefore not suitable for range-only measurements. Model-based encrypted control \cite{farokhiPrivacyDynamicalSystems2020} also introduces some privacy-preserving methods relevant to estimation. \cite{alexandruEncryptedCooperativeControl2019} proposes secure distributed control input aggregation using a novel weighted sum aggregation scheme, but requires repeated key distribution among all involved parties. In \cite{farokhiSecurePrivateControl2017}, a secure cloud-based control aggregator using partially homomorphic encryption is proposed, but is suitable only for untrusted service-providing cloud architectures.

While we have focused on the application of encryption schemes to localisation algorithms, homomorphic encryption schemes also require suitable number encoding when used with real numbers. As is the case with localisation sensors, real number measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. The Google BigNum library \cite{Encryptedbigqueryclient2015} provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In \cite{farokhiSecurePrivateControl2017}, a simpler alternative for encoding is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings.

@@

    Misbehaving sensors are a known problem in estimation theory, often requiring inconsistency or outlier detection mechanisms \cite{lazosSeRLocSecureRangeindependent2004,ben-galOutlierDetection2005}. As privacy-preserving implementations of such mechanisms require additional complications, we will not consider the scenario in this work.
    \item[Computational capabilities] The computational requirements for computing encryptions and homomorphic operations are typically more costly than their plaintext equivalents. We make the implicit assumption that all involved parties are computationally capable of computing the required encryption and filter operations.
\end{description}
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\footnotesize]
    % Sizes
    \tikzmath{\srange = 3.54; \sradius = 0.04;}
    % Sensor locations (sensor 3 is shared)
    \coordinate (s1) at (0.8523,-4.1508) {};
    \coordinate (s2) at (3.507,-5.5587) {};
    \coordinate (s3) at (2.5,-2.75) {} {};
    \coordinate (s4) at (4,-1.25) {};
    \coordinate (s5) at (5.4346,-3.1274) {};
    % Navigator location
    \coordinate (nav_loc) at (3.5,-3.25) {};
    % Bounding rectangle
    \coordinate (brect_start) at (0,0);
    \coordinate (brect_end) at (6.5, -6.5);

@@

% Include this if you use bibtex 
% and a bib file to produce the 
% bibliography (preferred). The
% correct style is generated by
% Elsevier at the time of printing.
\bibliographystyle{unsrt}
\bibliography{BibTeX/PrivateEIFLocalisation}
